---
title: "SVI Aggregate"
author: "Eric Stokan"
date: "6/7/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

# Set Working Directory

You will need to change to a working directory where you save the CSV files.

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/estokan/Desktop/Research/Data/Social Vulnerability Index")
```

# Libraries

The second thing you need to do is make sure you use these libraries.  They do require install if you have not used them.  Just type the library name in the *install packages* command install.packages('XXXXXXX').

```{r Libraries, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(dplyr)
library(stringr) #For dealing with string variables
library(lubridate) #For dealing with dates
```


# Data source

You will need to store these three svi data sources in your working directory so the R script can pull it in.

```{r Data SVI, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

SVI2000<- read.csv("SVI2000.csv")
SVI2010<-read.csv("SVI2010.csv")
SVI2014<-read.csv("SVI2014.csv")
```

# Merge data

Now, I need to merge all of the data together.

```{r Merge Data, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

SVI_All<- left_join(SVI2000, SVI2010, by=c("FIPS"))
SVI_All<- left_join(SVI_All, SVI2014, by=c("FIPS"))

```

# Subset data

Next, I really only want a few of the variables.  So, I subset the data and I change the variable names in the *select* command.  This has the same functionality as *select* and *rename*.

```{r SVI Calcualtions, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

SVI <-SVI_All %>%
  select(FIPS, GEOID=GEO_ID, STATE= STATE_NAME.x, STATEAbbr= STATE_ABBR.x, 
         County= COUNTY.x, T1_2000=USG1TP, T2_2000=USG2TP, T3_2000=USG3TP, 
         T4_2000=USG4TP, T_2000=USTP, T1_2010=R_PL_THEME1, T2_2010=R_PL_THEME2, 
         T3_2010=R_PL_THEME3, T4_2010=R_PL_THEME4, T_2010=R_PL_THEMES, 
         T1_2014=RPL_THEME1, T2_2014=RPL_THEME2, T3_2014=RPL_THEME3, 
         T4_2014=RPL_THEME4, T_2014=RPL_THEMES)
```

# Calculation variables

Next, I calculate all of the SVI measures

```{r SVI Mutations, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
SVI<-SVI %>%
  mutate(T1_Ch0010= (T1_2010-T1_2000)/T1_2000) %>%
  mutate(T2_Ch0010= (T2_2010-T2_2000)/T2_2000) %>%
  mutate(T3_Ch0010= (T3_2010-T3_2000)/T3_2000) %>%
  mutate(T4_Ch0010= (T4_2010-T4_2000)/T4_2000) %>%
  mutate(T_Ch0010= (T_2010-T_2000)/T_2000) %>%
  mutate(T1_Ch1014= (T1_2014-T1_2010)/T1_2010) %>%
  mutate(T2_Ch1014= (T2_2014-T2_2010)/T2_2010) %>%
  mutate(T3_Ch1014= (T3_2014-T3_2010)/T3_2010) %>%
  mutate(T4_Ch1014= (T4_2014-T4_2010)/T4_2010) %>%
  mutate(T_Ch1014= (T_2014-T_2010)/T_2010) %>%
  mutate(T1_Ch1400= (T1_2014-T1_2000)/T1_2000) %>%
  mutate(T2_Ch1400= (T2_2014-T2_2000)/T2_2000) %>%
  mutate(T3_Ch1400= (T3_2014-T3_2000)/T3_2000) %>%
  mutate(T4_Ch1400= (T4_2014-T4_2000)/T4_2000) %>%
  mutate(T_Ch1400= (T_2014-T_2000)/T_2000) 
```

```{r Write out data, echo=TRUE, message=FALSE, warning=FALSE}
#Write out the dataset
write.csv(SVI, file= "UpdatedSVI.csv")

```